Parsing Tweets into Universal Dependencies

### 2017-11-28 paper.tex
add speed comparison, add sota-stack comparison in pipeline eval. 2.3->2.6, remove incoherent ptb result.

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{green}{[#1 ---\textsc{Yi}]}}

@@

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets. Using the extended guidelines, we create
a new tweet treebank ({\sc Tweebank v2}), which has 55,607 tokens with
labeled attachments. 
It is more than four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreements
among our annotation and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.3 in LAS over \nascomment{which
  comparison is this?} the singlely trained baseline
and outperforms parsers that are state-of-the-art on other treebanks.
\end{abstract}

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank \citep[PTB]{Marcus93buildinga}.
Stanford dependencies were converted afterwards in their work  \nascomment{by who?}. 
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have  good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Based on these annotations, we designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.3 points \nascomment{again, not sure what this is quoting} in LAS over the baseline parser can be achieved with our
distillation method and that it outperforms other state-of-the-art parsers.

@@

\item We introduce and evaluate a system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  \nascomment{say it
  performs better than existing systems?}
\item We propose a new distillation
method for training a greedy parser, leading to better performance
than existing methods and without efficiency sacrifices.
\end{itemize}

@@

We report the inter-annotator agreement between the annotators
in the second stage. The agreement on POS is 96.61\%, the unlabeled
dependency agreement is 88.75\% and the labeled dependency agreement
is 84.31\%. Further analysis shows the major disagreements on POS
involve entity names (30.57\%) and topical hashtags (18.11\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{lr}
		\hline
		\it System & \it F1 \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\
% 		UD pipe v1.2 & 94.7 \\
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & \bf 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}

We trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training dataset
and tested on the \textsc{Tweebank v2} test set. 
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. \nascomment{we need to say what their model was;
  why does ours work better?  from their paper, it seems that they
  used a GRU instead of an LSTM ... is that all?  do we win because of
better tuning, or what?}
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & \it F1 \\
		\hline
		Stanford CoreNLP & 92.0 \\
		our tokenizer (\S\ref{sec:tok}) & 93.6 \\
		\hline
	\end{tabular}
	\caption{\citet{owoputi-EtAl:2013:NAACL-HLT} POS tagging performance with automatic tokenization on
          the \textsc{Tweebank v2} test set. \label{tbl:pos-result-vs-tok}}
\end{table}

@@

\citet{reimers-gurevych:2017:EMNLP2017} and others have
pointed out that neural network training is 
nondeterministic and depends on the seed for the random
number generator.
Our preliminary experiments confirm this finding, with a gap of 1 LAS on development data
between the best (75.8) and worst (74.8) runs. To control for this
effect, we report the average of five differently-seeded runs, for
each model and the compared one. %\nascomment{please check! we did this for all systems,
%  not just ours, right?}

%%%

@@

\paragraph{Initial results.}  The first section of Table~\ref{tbl:parse-result} compares
the stack LSTM with {\sc TweeboParser} (the system 
of \citealp{kong-EtAl:2014:EMNLP2014}) and the state-of-the-art parser
in the CoNLL 2017 evaluations, due to
 \citet{DBLP:journals/corr/DozatM16}, both of which are graph-based
 parsers requiring superlinear runtime. %\nascomment{can we quantify
%   how much time each one takes, either in tweets parsed per second or
%   seconds per tweet?}
Both of the comparison systems are re-trained on the same data as our
system.
Our system lags behind by nearly two LAS
points but runs faster than both of them.

\begin{table*}[t]
	\centering
	\begin{tabular}{rrcc}
		\hline
		\it Pipeline & Metrics & Our & SOTA stack\\
		\hline
		Tokenization & \it F1 & 98.3 & \\		
		POS tagging &  \it Precision F1 & 93.7 & \\
		Universal dependencies & \it LAS F1 & 74.1 & \\
		\hline
	\end{tabular}
	\caption{Evaluating our model in pipeline
		manner. \nascomment{should this be ``F1''?  shouldn't the
			last line be LAS?}\label{tbl:pipline} }
\end{table*}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}.
We see the
importance of segmentation:  without gold segmentation, UD parsing
performance drops by more than four points. We also compare
our model with a stack of the state-of-the-art system: Stanford Tokenizer,
ARK tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser. \yjcomment{tbd}

@@

\section{Conclusion}
We study the problem of parsing tweets into universal dependencies.
We adapt the UD guidelines to cover 
special constructions in tweets and create
the {\sc Tweebank v2}, which has 55,607 tokens. We characterize the disagreements
among our annotations and argue that inherent ambiguity in this genre
makes
consistent annotaiton a challenge.  Using this new treebank,
we build a pipeline system to parse tweets into UD. We also 
propose a new method to distill an ensemble of 20 greedy parsers into a single one
to overcome annotation noise without sacrificing efficiency.
Our parser achieves an improvement of 2.3  \nascomment{which
  comparison are you quoting here?} in LAS over a strong baseline
and outperforms other state-of-the-art parsers.

>>>>

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{green}{[#1 ---\textsc{Yi}]}}
%\newcommand{\yjcomment}[1]{}
%\newcommand{\nascomment}[1]{}
%\newcommand{\yicomment}[1]{}

@@

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets. Using the extended guidelines, we create
a new tweet treebank ({\sc Tweebank v2}), which has 55,607 tokens with
labeled attachments. 
It is more than four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreements
among our annotation and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over \nascomment{which
  comparison is this?}\yjcomment{my bad, it was fixed.} the baseline
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank \citep[PTB]{Marcus93buildinga}.
Stanford dependencies were converted afterwards in their work  \nascomment{by who?}\yjcomment{by themselves}. 
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have  good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Based on these annotations, we designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.6 points \nascomment{again, not sure what this is quoting}\yjcomment{fixed} in LAS over the
baseline parser can be achieved with our distillation method and that it outperforms
other state-of-the-art parsers in both accuracy and running speed.

@@

\item We introduce and evaluate a system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  Experimental results show
it performs better than a stack of the state-of-the-art alternatives.
\nascomment{say it
  performs better than existing systems?}\yjcomment{added.}

\item We propose a new distillation
method for training a greedy parser, leading to better performance
than existing methods and without efficiency sacrifices.
\end{itemize}

@@

We report the inter-annotator agreement between the annotators
in the second stage. The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & \it F1 \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\
% 		UD pipe v1.2 & 94.7 \\
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & \bf 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}

We trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training dataset
and tested on the \textsc{Tweebank v2} test set. 
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. \nascomment{we need to say what their model was;
  why does ours work better?  from their paper, it seems that they
  used a GRU instead of an LSTM ... is that all?  do we win because of
better tuning, or what?}\yjcomment{done.}
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & \it F1 \\
		\hline
		Stanford CoreNLP & 92.1 \\
		our bi-LSTM tokenizer (\S\ref{sec:tok}) & 93.6 \\
		\hline
	\end{tabular}
	\caption{\citet{owoputi-EtAl:2013:NAACL-HLT} POS tagging performance with automatic tokenization on
          the \textsc{Tweebank v2} test set. \label{tbl:pos-result-vs-tok}}
\end{table}

@@

\citet{reimers-gurevych:2017:EMNLP2017} and others have
pointed out that neural network training is 
nondeterministic and depends on the seed for the random
number generator.
Our preliminary experiments confirm this finding, with a gap of 1 LAS on development data
between the best (75.8) and worst (74.8) runs. To control for this
effect, we report the average of five differently-seeded runs, for
each of our models and the compared ones. \nascomment{please check! we did this for all systems,
  not just ours, right?}\yjcomment{yes. added}

@@

\paragraph{Initial results.}  The first section of Table~\ref{tbl:parse-result} compares
the stack LSTM with {\sc TweeboParser} (the system 
of \citealp{kong-EtAl:2014:EMNLP2014}) and the state-of-the-art parser
in the CoNLL 2017 evaluations, due to
 \citet{DBLP:journals/corr/DozatM16}, both of which are graph-based
 parsers requiring superlinear runtime. \nascomment{can we quantify
   how much time each one takes, either in tweets parsed per second or
   seconds per tweet?}\yjcomment{done.}
Both of the comparison systems are re-trained on the same data as our
system.
Our system lags behind by nearly two LAS
points but runs faster than both of them.

\begin{table*}[t]
	\centering
	\begin{tabular}{rrcc}
		\hline
		\it Pipeline & Metrics & ours & SOTA stack\\
		\hline
		Tokenization & \it F1 & 98.3 & 96.6 \\		
		POS tagging &  \it Precision F1 & 93.7 & 92.1 \\
		Universal dependencies & \it LAS F1 & 74.1 & 70.3 \\
		\hline
	\end{tabular}
	\caption{Evaluating our model in pipeline
		manner. \nascomment{should this be ``F1''?  shouldn't the
			last line be LAS?}\yjcomment{added metrics column}\label{tbl:pipline} }
\end{table*}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a stack of the state-of-the-art systems (named as the {\it SOTA stack}):
Stanford CoreNLP tokenizer,
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser.
Our system differs with the SOTA stack in the tokenization and parser components.
From Table \ref{tbl:pipline}, our system outperforms the SOTA stack when
evaluated in pipeline manner.
The results also emphasize the
importance of segmentation:  without gold segmentation, UD parsing
performance drops by more than four points.

@@

\section{Conclusion}
We study the problem of parsing tweets into universal dependencies.
We adapt the UD guidelines to cover 
special constructions in tweets and create
the {\sc Tweebank v2}, which has 55,607 tokens. We characterize the disagreements
among our annotations and argue that inherent ambiguity in this genre
makes
consistent annotaiton a challenge.  Using this new treebank,
we build a pipeline system to parse tweets into UD. We also 
propose a new method to distill an ensemble of 20 greedy parsers into a single one
to overcome annotation noise without sacrificing efficiency.
Our parser achieves an improvement of 2.6  \nascomment{which
  comparison are you quoting here?}\yjcomment{fixed} in LAS over a strong baseline
and outperforms other state-of-the-art parsers in both accuracy and speed.
